Naturalness in U(1) extensions of the MSSM

### 2015-01-21 paper/e6ssm-tuning-paper.tex
Added reference for Z' reach of LHC run II
added title/abstract options in comments.
added missing 'we'

\begin{abstract}
The abstract.
% Copied SUSY2014 abstract as placeholder
% With the discovery of a 125 GeV Standard Model (SM) like Higgs boson
% and rising lower bounds on the masses of superpartners, minimal models
% of low energy supersymmetry (SUSY) appear to be increasingly fine
% tuned. In the Minimal Supersymmetric Standard Model (MSSM) large stop
% masses, needed to obtain the observed Higgs mass, give a significant
% contribution to fine tuning. $U(1)$ extensions of the MSSM that solve
% the $\mu$-problem have new F- and D-term contributions to the Higgs
% mass, allowing a 125 GeV Higgs with lighter stops and so alleviating
% the associated fine tuning. However, it was recently demonstrated,
% within a constrained version of the Exceptional Supersymmetric
% Standard Model (E$_6$SSM) defined at the Grand Unification (GUT)
% scale, that the presence of a massive $Z'$ boson introduces a
% substantial new source of tuning in these theories at tree level. Here
% we consider the impact of this tuning when we remove the fine tuning
% that depends on assumptions about the SUSY breaking scale, by setting
% the SUSY breaking parameters at low energies. We find that the
% existing experimental bounds on the $Z'$ mass impose an effective
% lower bound on the fine tuning, which does not depend on assumptions
% about SUSY breaking.  We compare our results against the tuning in the
% MSSM with the SUSY breaking terms also defined at low energies.

@@

While this approach has many merits here will employ the traditional
measure of fine tuning because it is both simple to apply and easy to
compare with previous results due to it's widespread use.  Fortunately
the derivatives which appear in these tuning are also similar to the
Bayesian motivated measure so there should not be too large a
discrepancy between the two approaches.

@@

Since the stop mass does not have such a large impact on the fine
tuning when the cutoff scale is very low we can use this to see more
clearly the impact of the $Z^\prime$ mass on fine tuning. To do so we
select a fixed low cutoff of $M_X=20$ TeV and compare the fine tuning
for two different values of the $Z^\prime$ mass. We choose to look at
$M_{Z^\prime}=2.5$ TeV, which is just above the current limits, and
$M_{Z^\prime}=4.5$ TeV, which should be in reach in run II at the LHC
\cite{}[CITE PLEASE] and then compare the fine tuning calculated in
each case to the tuning in the MSSM.  For this, we have performed a
large six dimensional parameter space scan in both the MSSM and
E$_6$SSM, varying those parameters most relevant for the fine tuning
and the Higgs mass.  Therefore the set of parameters which we vary
includes $\mu$, $B$ and $\tan \beta$ for the MSSM, and $\lambda$,
$A_\lambda$ and $\tan \beta$, for the E$_6$SSM, which appear at tree
level in the EWSB conditions of the models. While the RGE contribution
from large stop masses to the fine tuning is small for such a low
cut-off scale, the stop contributions to the effective potential can
play a significant role in reducing the fine tuning [WE SHOULD DISCUSS
  THIS].  For this reason it is still important to properly treat the
tuning associated with stop contributions to the one-loop effective
potential, and so we also scan over the soft masses $m_{Q_3}^2$,
$m_{u_3}^2$ and the stop mixing $A_t$.  The relevant parameters and
ranges that were scanned over are summarised in Table
\ref{tab:scanranges}. With the exception of the stop mixing $A_t$ and
the soft trilinear $A_\lambda$, which are logarithmically scanned, all
parameters are linearly scanned.  In addition to this we also repeat
each scan for three different values of $M_2$ to allow more variation
in the chargino masses.

>>>>

\begin{abstract}
The abstract.
%%Original abstract
% Copied SUSY2014 abstract.
%% With the discovery of a 125 GeV Standard Model (SM) like Higgs boson
%% and rising lower bounds on the masses of superpartners, minimal models
%% of low energy supersymmetry (SUSY) appear to be increasingly fine
%% tuned. In the Minimal Supersymmetric Standard Model (MSSM) large stop
%% masses, needed to obtain the observed Higgs mass, give a significant
%% contribution to fine tuning. $U(1)$ extensions of the MSSM that solve
%% the $\mu$-problem have new F- and D-term contributions to the Higgs
%% mass, allowing a 125 GeV Higgs with lighter stops and so alleviating
%% the associated fine tuning. However, it was recently demonstrated,
%% within a constrained version of the Exceptional Supersymmetric
%% Standard Model (E$_6$SSM) defined at the Grand Unification (GUT)
%% scale, that the presence of a massive $Z'$ boson introduces a
%% substantial new source of tuning in these theories at tree level. Here
%% we consider the impact of this tuning when we remove the fine tuning
%% that depends on assumptions about the SUSY breaking scale, by setting
%% the SUSY breaking parameters at low energies. We find that the
%% existing experimental bounds on the $Z'$ mass impose an effective
%% lower bound on the fine tuning, which does not depend on assumptions
%% about SUSY breaking.  We compare our results against the tuning in the
%% MSSM with the SUSY breaking terms also defined at low energies.

%%Alternative 1
%% Alt title: Probing natural SUSY in the most general way in the MSSM and U(1) extensions.

%% The discovery of a 125 GeV Standard Model (SM) like Higgs boson and
%% rising lower bounds on the masses of superpartners, have lead to
%% concerns that supersymmetric models are now fine tuned.  In the
%% Minimal Supersymmetric Standard Model (MSSM) large stop masses,
%% required for a $125$ GeV Higgs, also feed into the EWSB conditions
%% through renormalisation group equations forcing one to fine tune these
%% parameters to obtain the correct electroweak vacuum expectation
%% value. Nonetheless this fine tuning depends crucially on our
%% assumptions about the SUSY breaking scale.  Here we consider what fine
%% tuning may appear in supersymmetry at tree level, so that it is not so
%% dependent on assumptions about the SUSY breaking mechanism.  We focus
%% on U(1) extensions since these provide the most elegant solution to
%% the $\mu$-problem, which is also a naturalness issue, but we also
%% compare to the simplest SUSY model, the MSSM. In U(1) extensions anew
%% F- and D-term contributions raise the Higgs mass, allowing a $125$ GeV
%% Higgs with lighter stops which can alleviate this stop fine tuning in
%% models with SUSY breaking fixed at the high scale. However the
%% presence of a massive $Z'$ boson introduces a substantial new source
%% of tuning in these theories at tree level.  By choosing a low scale
%% boundary condition for where the SUSY breaking parameters are set we
%% compare how this tree level tuning compares to that of the MSSM where
%% chargino mass limits are the best way to probe tree level fine tuning,
%% by raising $\mu$ directly.  We find that the $Z^\tprime$ tuning places
%% more stringent limits on the fine tuning in most of the $U(1)$
%% extensions and that increasing $Z^\prime$ limits will challenge the
%% most natural scenarios of these models.

%% Alternative 2

%% Alt title: Z' mass limits and the naturalness of supersymmetry

%% The discovery of a 125 GeV Higgs boson and rising lower bounds on the
%% masses of superpartners, have lead to concerns that supersymmetric
%% models are now fine tuned.  Large stop masses, required for a $125$
%% GeV Higgs, feed into the EWSB conditions through renormalisation group
%% equations forcing one to fine tune these parameters to obtain the
%% correct electroweak vacuum expectation value.  Nonetheless this fine
%% tuning depends crucially on our assumptions about the SUSY breaking
%% scale.  At the same time U(1) extensions provide the most compelling
%% solution to the $\mu$-problem, which is also a naturalness issue.
%% These very well motovated supersymmetric models predict a new
%% $Z^\prime$ boson which could be discovered at the LHC.  like the
%% superpartners naturalness of the model requires that the $Z^\prime$
%% boson mass should not be too \far above the TeV scale.  Moreover this
%% fine tuning appears at the tree level, making it less dependent on a
%% SUSY breaking mecahnism which has not been fully determnined.  Here we
%% study this fine tuning for several $U(1)$ supesymmetric extensions of
%% the standard model and compare it to the situation in the MSSM where
%% the most direct tree level fine tuning can be probed through chargino
%% mass limits.  we show that future LHC $Z^\prime$ searches are
%% extremely important for challenging the most natural scenarios in
%% these models.

@@

While this approach has many merits here we will employ the traditional
measure of fine tuning because it is both simple to apply and easy to
compare with previous results due to it's widespread use.  Fortunately
the derivatives which appear in these tuning are also similar to the
Bayesian motivated measure so there should not be too large a
discrepancy between the two approaches.

@@

Since the stop mass does not have such a large impact on the fine
tuning when the cutoff scale is very low we can use this to see more
clearly the impact of the $Z^\prime$ mass on fine tuning. To do so we
select a fixed low cutoff of $M_X=20$ TeV and compare the fine tuning
for two different values of the $Z^\prime$ mass. We choose to look at
$M_{Z^\prime}=2.5$ TeV, which is just above the current limits, and
$M_{Z^\prime}=4.5$ TeV, which should be in reach in run II at the LHC
\cite{Godfrey:2013eta} and then compare the fine tuning calculated in
each case to the tuning in the MSSM.  For this, we have performed a
large six dimensional parameter space scan in both the MSSM and
E$_6$SSM, varying those parameters most relevant for the fine tuning
and the Higgs mass.  Therefore the set of parameters which we vary
includes $\mu$, $B$ and $\tan \beta$ for the MSSM, and $\lambda$,
$A_\lambda$ and $\tan \beta$, for the E$_6$SSM, which appear at tree
level in the EWSB conditions of the models. While the RGE contribution
from large stop masses to the fine tuning is small for such a low
cut-off scale, the stop contributions to the effective potential can
play a significant role in reducing the fine tuning [WE SHOULD DISCUSS
  THIS].  For this reason it is still important to properly treat the
tuning associated with stop contributions to the one-loop effective
potential, and so we also scan over the soft masses $m_{Q_3}^2$,
$m_{u_3}^2$ and the stop mixing $A_t$.  The relevant parameters and
ranges that were scanned over are summarised in Table
\ref{tab:scanranges}. With the exception of the stop mixing $A_t$ and
the soft trilinear $A_\lambda$, which are logarithmically scanned, all
parameters are linearly scanned.  In addition to this we also repeat
each scan for three different values of $M_2$ to allow more variation
in the chargino masses.
